---
title: "Human Activity Recognition"
author: "moocdoer"
date: "26 October 2014"
output: html_document
---

## Executive Summary

The objective of this project is predict human activities (classified A to E) by device data.

## Data loading

The train set will be used as the data set of this project. The test set made available does not include the 'classe' variable.

```{r load_data}
# Load the dataset
dataset = read.csv('pml-training.csv')
```

The dataset was reviewed previously and several variables are not needed or fully NA. For this reason I will use only the device data that has numerical properties.

```{r subset_data}
# Load dplyr
library(dplyr)

# Select only the wanted columns
dataset = select(
    dataset,
    roll_belt,pitch_belt,
    yaw_belt,
    total_accel_belt,
    gyros_belt_x,
    gyros_belt_y,
    gyros_belt_z,
    accel_belt_x,
    accel_belt_y,
    accel_belt_z,
    magnet_belt_x,
    magnet_belt_y,
    magnet_belt_z,
    roll_arm,
    pitch_arm,
    yaw_arm,
    total_accel_arm,
    gyros_arm_x,
    gyros_arm_y,
    gyros_arm_z,
    accel_arm_x,
    accel_arm_y,
    accel_arm_z,
    magnet_arm_x,
    magnet_arm_y,
    magnet_arm_z,
    roll_dumbbell,
    pitch_dumbbell,
    yaw_dumbbell,
    total_accel_dumbbell,
    gyros_dumbbell_x,
    gyros_dumbbell_y,
    gyros_dumbbell_z,
    accel_dumbbell_x,
    accel_dumbbell_y,
    accel_dumbbell_z,
    magnet_dumbbell_x,
    magnet_dumbbell_y,
    magnet_dumbbell_z,
    roll_forearm,
    pitch_forearm,
    yaw_forearm,
    total_accel_forearm,
    gyros_forearm_x,
    gyros_forearm_y,
    gyros_forearm_z,
    accel_forearm_x,
    accel_forearm_y,
    accel_forearm_z,
    magnet_forearm_x,
    magnet_forearm_y,
    magnet_forearm_z,
    classe
)
```

## Prediction Model

Data are divided into train (80%) and test (20%) sets. These will be used to train a random forest and measure the classification error.

```{r fit_model}
# Load libs
library(caret)
library(randomForest)

# Set seed
set.seed(1)

# Create train and test data
inTrain = createDataPartition(y = dataset$classe, p = 0.8, list = F)
trainSet = dataset[inTrain, ]
testSet = dataset[-inTrain, ]

# Fit a random forest
modelFit = randomForest(classe ~ ., data = trainSet, p = .1, ntree = 100)

# Training classification error and confusion matrix
trainPred = predict(modelFit, newdata = trainSet)
sum(trainPred != trainSet$classe) / length(trainPred)
table(trainPred, trainSet$classe)

# Testing classification error and confusion matrix
testPred = predict(modelFit, newdata = testSet)
sum(testPred != testSet$classe) / length(testPred)
table(testPred, testSet$classe)
```

Training classification error is 0%.
Testing classification error is 0.7%